\chapter{Bullet points \& arrangement of the summary}
\setlength{\headheight}{12.71342pt}
\addtolength{\topmargin}{-0.71342pt}

\begin{itemize}
    \item \textbf{Title:}  
    \subitem \textit{Membrane-Associated Enzymes in Milk: Classification, Function, and Recent Insights}

    \item \textbf{1. Introduction}  
    \subitem Overview of membrane-associated enzymes in milk.  
    \subitem Importance in dairy science and milk composition.  
    \subitem Scope of the summary (focus on sulfhydryl oxidase, xanthine oxidoreductase, $\gamma$-glutamyltransferase, and 5'-nucleotidase).  

    \item \textbf{2. Classification and Mechanisms of Enzymes}  
    \subitem \textbf{Sulfhydryl Oxidase} – Catalyzes oxidation of protein thiols, disulfide bond formation, role in protein folding.  
    \subitem \textbf{Xanthine Oxidoreductase (XOR)} – Converts hypoxanthine to xanthine and uric acid, exists as xanthine dehydrogenase or oxidase.  
    \subitem \textbf{$\gamma$-Glutamyltransferase (GGT)} – Transfers $\gamma$-glutamyl groups, involved in glutathione metabolism and amino acid transport.  
    \subitem \textbf{5'-Nucleotidase} – Hydrolyzes 5'-mononucleotides, contributes to nucleotide metabolism.  

    \item \textbf{3. Role of Enzymes in Milk and Dairy Products}  
    \subitem \textbf{Sulfhydryl Oxidase} – Stabilization of milk proteins, impact on UHT milk flavor.  
    \subitem \textbf{XOR} – Antimicrobial properties, role in milk fat globule secretion.  
    \subitem \textbf{GGT} – Supports amino acid availability for milk protein synthesis, possible thermal processing marker.  
    \subitem \textbf{5’-Nucleotidase} – Potential role in neonatal digestion and milk secretion studies.  

    \item \textbf{4. Summary of Recent Scientific Studies}  
    \subitem \textbf{Sulfhydryl Oxidase} – Research on enzymatic activity and protein cross-linking effects.  
    \subitem \textbf{XOR} – Studies on antimicrobial activity and oxidative rancidity.  
    \subitem \textbf{GGT} – Investigations into its use as a marker for milk pasteurization.  
    \subitem \textbf{5’-Nucleotidase} – Research on its extraction and enzymatic function in dairy processing.  

    \item \textbf{5. Conclusion}  
    \subitem Key takeaways on membrane-associated enzymes in milk.  
    \subitem Significance for dairy processing, product quality, and nutrition.  
    \subitem Future research directions.  

    \item \textbf{6. References}  
    \subitem Full bibliographic information in Harvard (author-year) style, formatted using \texttt{natbib}.  
\end{itemize}


