\chapter{Group Project - Enzymes - Membrane Associated Enzymes}
\setlength{\headheight}{12.71342pt}
\addtolength{\topmargin}{-0.71342pt}

\subsection*{Formal requirements for the summary}

\begin{itemize}
    \item The summary should not be longer than 3-4 pages excluding references
    \item The summary should be written as continuous text (a small number of bullet points may be used)
    \item All references must be cited with full bibliographic information. Citation style can be chosen individually, but should be consistent throughout the summary
    \item Please address the following aspects:
    \subitem An appropriate title and suitable headings
    \subitem Classification and modus operandi of the enzyme(s)
    \subitem The role of the enzyme(s) in milk and/or dairy products
    \subitem A short summary of 1-2 recent scientific studies related to the enzyme(s)
    \subitem References
    \item The deadline for submission is February 28
\end{itemize}

\section{Introduction}
\begin{itemize}
    \item Phospholipids (PL) in Milk: Estimates of PL content in bovine milk vary (0.9–2.3\% of total lipids) depending on extraction methods (Röse Gottlieb vs. Folch extraction).
    \item Milk Fat Globule Membrane (MFGM): Contains 60–65\% of milk PL, with the remaining 35–40\% found in the skim milk phase.
    \item Extracellular Vesicles (EVs): Recently recognized nano-sized phospholipid structures in skim milk, distinct from milk fat globules, but involved in intercellular communication.
    \item Membrane-Associated Enzymes: Identified in both MFGs and EVs, but their abundance is generally low. Many originate from ER, Golgi, or cytosolic crescents.
    \item Activity Considerations: Many of these enzymes remain inactive in milk due to the absence of substrates or unsuitable environmental conditions.
    \item Scope of Discussion: The summary will focus on enzymes relevant to mammary gland biology, milk integrity, and physiological functions upon consumption, excluding those related to lipid synthesis.
\end{itemize}

\section{Sulfhydryl Oxidase}


\subsection*{Structure of Flavin-Dependent Sulfhydryl Oxidase}


\subsection*{Biological Role of Flavin-Dependent Sulfhydryl Oxidase and Significance in Milk}

\cite*{RM_01}

\section{Catalse}


\subsection*{Structure of Catalase}


\subsection*{Biological Role of Catalase and Significance in Milk}


\section{Lactoperoxidase}


\subsection*{Structure of Lactoperoxidase}


\subsection*{Biological Role of Lactoperoxidase and Significance in Milk}


\section{Xanthine Oxidoreductase}


\subsection*{Structure of Xanthine Oxidoreductase}



\subsection*{Biological Role of Xanthine Oxidoreductase and Significance in Milk}

 
